---
title: "Example"
output:
  html_document:
    df_print: paged
---

This is the notebook to illustrate the usage of the `ZCAPM` package. 

## List of existing functions and dataset in the package 
```{r}
library(ZCAPM)
ls("package:ZCAPM")
```

* `ff25_day`, `ind47`, `ff_factors`, and `mu_sigma` are the four datasets we included in this package as examples. The first two are datasets for 25 size-B/M (book-to-market equity) sorted portfolios and 47 industry portfolios from January 1964 to December 2015, while the rest two store the factors can be considered in the model. 

* `GetMonthReturn` can be used to get monthly return according to daily return. 

* `EMRegression`, `estLM` and `FactorModel` are three functions used for traditional models, e.g. CAPM market model, Fama and French's three-factor model, Carhart's four-factor model, and Fama and French's five-factor model. You can design your own factor model as well.

* `EM_loop`, `estZCAPM` and `ZCAPM` are three functions used for our proposed ZCAPM with EM algorithm.

* `SingleReg` and `fm_test` are the functions used for the two-step Fama and MacBeth (1973) procedure for cross-sectional tests. 

* `MonthRoll` and `RollForward` are the functions for month-rolling estimation and apply a certain function to a data in a rolling window style.

For more detailed information, please use `help()` for each function of interest.


## Example for illustration

### Load the dataset and data preprocessing
```{r}
data("SizedBM25") # 25 size B/M portfolios
data("mu_sigma") 
data("ff_factors")

# select portfolio excess return for period of interest
# get factor data for period of interest
start.date <- 19640101
end.date <- 20151231
data.port <- subset(ff25_day, Date >= start.date & Date <= end.date)
data.factor <- subset(mu_sigma, Date >= start.date & Date <= end.date)
data.factor1 <- subset(ff_factors, Date >= start.date & Date <= end.date)
data.factor$SMB <- data.factor1$SMB
data.factor$HML <- data.factor1$HML
rm(data.factor1)

# check if the dates of portfolios and covariates matched
stopifnot(all(data.port$Date == data.factor$Date))

# get year and month
data.yyyymm <- as.numeric(substr(data.port$Date, 1, 6))
month.list <- sort(unique(data.yyyymm))

# get excess return of each portfolio
data.port <- data.port[, -1]
data.port.exc.ret <- as.data.frame(apply(data.port, 2, FUN = function(x, y){x - y}, data.factor$R_f))

# get monthly excess return of each portfolio
port.exc.ret.month <- aggregate(data.port.exc.ret, by = list(yyyymm = data.yyyymm), FUN = GetMonthReturn)[, -1]
```

### Set the parameters 
```{r}
# set rolling window, unit: month
roll.width <- 12
tol <- 0.001
MaxIter <- 1000

# criterion = 1: if 'abs(diff(new - old)/old)' is used as convergence criterion
# criterion = 2: if 'abs(diff(new - old)/(abs(old) + 1))' is used as the convergence criterion
# criterion = 2 usually leads to better convergence
criterion <- 1  
params <- c(tol, MaxIter, criterion)

# keep only the one-month ahead monthly return in testing period
port.exc.ret.month.test <- port.exc.ret.month[-(1:roll.width), ]

# Specify the months used for model estimation
month.list.est <- month.list[-length(month.list)]
```

### Estimation using factor model and ZCAPM
```{r}
# zcapm model
zcapm.res <- ZCAPM(data.port.exc.ret, month.list.est, roll.width, data.yyyymm, data.factor[, c("R_a.R_f", "sigma_a")], params)

# capm and 3-factor model
capm.res <- FactorModel(data.port.exc.ret, month.list.est, roll.width, data.yyyymm, data.frame(R_a.R_f = data.factor[, "R_a.R_f"]))
ff3.res <- FactorModel(data.port.exc.ret, month.list.est, roll.width, data.yyyymm, data.factor[, c("R_a.R_f", "SMB", "HML")])

# adjustment to the zeta's of zcapm
num.mod <- nrow(zcapm.res$factor.loading) / 2
zcapm.res$factor.loading[(num.mod + 1): (num.mod * 2), ] <- zcapm.res$factor.loading[(num.mod + 1): (num.mod * 2), ] * 21
```

### Out-of-sample cross-sectional Fama and MacBeth test
```{r}
# Specify the list of models to be tested
# The name of variables stand for the models, this will be the name of model show in the summary table
fact.load.lst <- list(ZCAPM = zcapm.res,
                      LM = capm.res,
                      FF3 = ff3.res)

# run the FM-test
test_res <- fm_test(port.exc.ret.month.test, fact.load.lst)

# print out the summary table
print.summary(test_res)
```
